---
title: "Linguistic Predictors of Concreteness"
author: "Bodo Winter"
date: "3/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This script analyzes several linguistic predictors of concreteness and abstractness:

* part-of-speech
* number of morphemes
* count/mass distinction
* etymology

First, each linguistic factor will be analyzed separately; then they will be analyzed in a simultaneous regression analysis.

Load in the data:

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(car) # for variance inflation factors vif()
library(broom) # for glance()
library(rsq)
library(gridExtra)
library(textstem) # for lemmatize_strings()
library(effsize)

# Brysbaert concreteness ratings:

conc <- read_csv('../data/brysbaert_concreteness.csv')

# SUBTLEX POS Tags:

SUBTL <- read_csv('../data/SUBTLEX_US_with_POS.csv')

# ELP POS tags & ELP number of morphemes:

ELP <- read_csv('../data/ELP_with_POS_cleaned.csv')

# BECL countability data:

BECL <- read_csv('../data/BECL.csv')

# Etymologies:

etym <- read_csv('../data/monaghan_roberts_etymologies.csv')

# Load in morpholex data:

morph <- read_csv('../data/morpholex_one_derivational_suffix.csv')
nosuffix <- read_csv('../data/morpholex_no_suffix.csv')

# Timo's beautiful ggplot2 theme:

source('theme_timo.R')
```

## Preprocessing:

Lemmatize all strings (get rid of inflectional morphology), except for -est, which we treat as derivational here given that's how it is treated in MorphoLex (see paper for discussion of -est).

```{r}
conc$Lemmas <- conc$Word
ids <- !str_detect(conc$Word, 'est')
conc[ids, ]$Lemmas <- lemmatize_strings(conc[ids, ]$Word)

# Get rid of 'duplicates'

conc <- filter(conc,
               !duplicated(Lemmas))

# Replace "Word" column with "lemma" column:

conc <- mutate(conc,
               Word = Lemmas)
```

Merge SUBTLEX POS and ELP data into conc:

```{r}
# SUBTLEX parts-of-speech tags:
conc$SUBTL_POS <- SUBTL[match(conc$Word, SUBTL$Word), ]$Dom_PoS_SUBTLEX
conc$SUBTL_percent <- SUBTL[match(conc$Word, SUBTL$Word), ]$Percentage_dom_PoS
conc$SUBTL_allPOS <- SUBTL[match(conc$Word, SUBTL$Word), ]$All_PoS_SUBTLEX

# ELP parts-of-speech tags:

conc$ELP_POS <- ELP[match(conc$Word, ELP$Word), ]$POS

# Number of letters from ELP:

conc$Letters <- ELP[match(conc$Word, ELP$Word), ]$Length

# Number of phonemes:

conc$NPhon <- ELP[match(conc$Word, ELP$Word), ]$NPhon

# Number of morphemes:

conc$NMorph <- ELP[match(conc$Word, ELP$Word), ]$NMorph

# Number of morphemes:

conc$NMorph <- ELP[match(conc$Word, ELP$Word), ]$NMorph
```

Process etymologies... first clean the strings:

```{r}
etym <- filter(etym, word != '#NAME?')
etym <- mutate(etym,
               word = str_replace(word, '\\(.+\\)', ''),
               word = str_trim(word),
               word = str_to_lower(word))
```

Next, sort the labels:

```{r}
etym$Last <- str_split(etym$links, '%', simplify = TRUE)[, 1]
```

Match this into the concreteness file:

```{r}
conc$Etym <- etym[match(conc$Word, etym$word), ]$Last

conc_counts <- conc %>% count(Etym, sort = TRUE) %>%
  mutate(large = ifelse(n > 100, 'large', 'small')) %>%
  filter(!is.na(Etym))

conc$EtymSize <- conc_counts[match(conc$Etym, conc_counts$Etym), ]$large
```

Merge the following categories:

- Proto-Germanic / Old English / Middle English = English/Germanic
- Late Latin / Latin / Medieval Latin = Latin
- Middle French / Old French / French / Anglo-French = French

```{r}
french <- c('Anglo-French', 'French', 'Middle French', 'Old French')
latin <- c('Late Latin', 'Latin', 'Medieval Latin', 'Modern Latin')
english <- c('Middle English', 'Old English', 'Proto-Germanic')
other <- c('Greek', 'Italian', 'Old Norse')
conc$EtymSimplified <- NA
conc <- mutate(conc,
               EtymSimplified = ifelse(Etym %in% french, 'French', EtymSimplified),
               EtymSimplified = ifelse(Etym %in% latin,
                                       'Latin', EtymSimplified),
               EtymSimplified = ifelse(Etym %in% english,
                                       'English', EtymSimplified),
               EtymSimplified = ifelse(Etym %in% other, 'Other', EtymSimplified))
```

In a further step — this was requested by Francesca & Marianna, we will merge "Other" and "English", as the hypothesis is about French-based versus the rest.

```{r}
conc <- mutate(conc,
               EtymSimplified = ifelse(EtymSimplified == 'English',
                                       'Other', EtymSimplified))
```

Create a new 'function word' category:

```{r}
function_words <- c('Conjunction', 'Determiner',
                    'Preposition', 'Pronoun', 'Article',
                    'Ex', 'To', 'Not')
conc <- mutate(conc,
               SUBTL_POS = ifelse(SUBTL_POS %in% function_words,
                                  'Function', SUBTL_POS))
rm(function_words)
```

Set everything else NA:

```{r}
conc <- mutate(conc,
               SUBTL_POS = ifelse(SUBTL_POS == '#N/A', NA, SUBTL_POS),
               SUBTL_POS = ifelse(SUBTL_POS == 'Interjection', NA, SUBTL_POS),
               SUBTL_POS = ifelse(SUBTL_POS == 'Letter', NA, SUBTL_POS),
               SUBTL_POS = ifelse(SUBTL_POS == 'Name', NA, SUBTL_POS),
               SUBTL_POS = ifelse(SUBTL_POS == 'Number', NA, SUBTL_POS),
               SUBTL_POS = ifelse(SUBTL_POS == 'Unclassified', NA, SUBTL_POS))
```

Add morpheme parses (suffix info):

```{r}
conc$MorphParse <- morph[match(conc$Word, morph$Word), ]$MorphoLexSegm
```

Create a table of the morphemes:

```{r}
these_morphs <- c('>ly', '>y', '>er', '>ion', '>al', '>ness', '>ic',
                  '>ate', '>able', '>est', '>ious', '>ity', '>ive',
                  '>ant', '>ist', '>ize', '>less', '>ory', '>ory',
                  '>ful', '>ance')
suffixes <- c('-ly', '-y', '-er', '-ion', '-al', '-ness', '-ic',
              '-ate', '-able', '-est', '-ious', '-ity', '-ive',
              '-ant', '-ist', '-ize', '-less', '-ory', '-ory',
              '-ful', '-ance')
```

Attach these to the concreteness dataset:

```{r}
conc$Suffix <- NA
conc$Suffix <- as.character(conc$Suffix)

for (i in 1:length(these_morphs)) {
  this_morph <- these_morphs[i]
  conc[which(str_detect(conc$MorphParse, this_morph)), ]$Suffix <- suffixes[i]
}
```

Check:

```{r}
table(conc$Suffix)
```

Check words that have these suffixes versus words that do not.

```{r}
conc <- mutate(conc,
               HasSuffix = ifelse(!is.na(Suffix), 'has suffix', 'no suffix'))
```

Since the 'suffix' parse is only for bimorphemic words, the 'no suffix' category needs to be set to NA for NMorph > 2.

```{r}
conc[which(conc$NMorph > 2), ]$HasSuffix <- NA
```

Create a "Suffix" column that also includes the no suffixes for plotting purposes:

```{r}
conc <- mutate(conc,
               SuffixWithMono = ifelse(is.na(Suffix), 'monomorphemic', Suffix))
conc[which(conc$NMorph > 1 & is.na(conc$Suffix)), ]$SuffixWithMono <- NA
conc[which(is.na(conc$NMorph) & is.na(conc$Suffix)), ]$SuffixWithMono <- NA
```

Are standard deviations influenced by word knowledge?

```{r}
summary(lm(Conc.SD ~ Percent_known, data = conc))
```

Yes, but not massively so.

Exclude words that were known by less than 95% of all people in the concreteness rating study. This makes sense because these data are probably noisy:

```{r}
# How much:

sum(conc$Percent_known < 0.95) / nrow(conc)
sum(conc$Percent_known < 0.95)

# Exclude:

conc <- filter(conc, Percent_known > 0.95)

```

A whopping 21% are excluded... or alternatively, 5820 data points.


Get all compounds:

```{r, message = FALSE, warning = FALSE}
juhasz <- read_csv('../data/juhasz_lai_woodcock_2015.csv')
LADEC <- read_csv('../data/LADECv1-2019.csv')
kim <- read_csv('../data/kim_yap_goh_2019.csv')
```

Get all compound words into one big list:

```{r}
compounds <- c(juhasz$Compound, LADEC$stim, kim$CW)
compounds <- unique(compounds)
```

Add a "is compound" variable to the conc data frame:

```{r}
conc <- mutate(conc,
               IsCompound = ifelse(Word %in% compounds, 'Compound', 'No compound'))
```

Crete a subset with low SD items. Here I use an 50th percentile cut-off (median split), so the words with the 50% lowest SD. Other values would make sense too (feel free to change them) and don't affect the results vastly:

```{r}
conc_sub <- filter(conc,
                   Conc.SD < quantile(Conc.SD, 0.5))
nrow(conc_sub)
nrow(conc) - nrow(conc_sub)
1 - (nrow(conc_sub) / nrow(conc))
```

This excludes 11,481 data points.

The analyses will be done with both datasets (all SDs allowed and low SD).

## Linguistic factor 1: Part of speech:

Only look at nouns, verbs and adjectives, SUBTLEX:

```{r, fig.width = 10, fig.height = 6}
POS <- c('Verb', 'Adjective', 'Noun', 'Adverb', 'Function')
conc_red <- filter(conc,
                   SUBTL_POS %in% POS)

# Plot and aesthetics:

conc_POS <- conc_red %>%
  mutate(SUBTL_POS = factor(SUBTL_POS,
                            levels = c('Adverb', 'Function', 'Adjective',
                                       'Verb', 'Noun'))) %>% 
  ggplot(aes(x = SUBTL_POS, y = Conc.M, fill = SUBTL_POS))

# Geoms:

conc_POS <- conc_POS + geom_boxplot()

# Themes and other properties:

conc_POS <- conc_POS +
  labs(x = '', y = 'Concreteness\n') +
  theme_timo +
  theme(legend.position = 'none') + 
  scale_fill_brewer(palette = 'Accent')

# Look and save:

conc_POS
ggsave(plot = conc_POS, filename = '../figures/POS.pdf',
       width = 8, height = 4.5)
```

Get the descriptive values for reporting:

```{r}
conc_red %>% group_by(SUBTL_POS) %>%
  summarize(M = mean(Conc.M),
            SD = sd(Conc.M),
            SD = round(SD, 2),
            N = n()) %>% 
  arrange(desc(M))
```

Nouns are a whopping 1 concreteness pint above adjectives, and about half a point above verbs.

Make a model of this:

```{r}
xmdl <- lm(Conc.M ~ SUBTL_POS, conc_red)
anova(xmdl)
summary(xmdl)$r.squared
```

There is a significant effect of parts-of-speech, and it explains 19% of the variance in concreteness ratings!

Repeat for low SD words:

```{r}
conc_red <- filter(conc_sub,
                   SUBTL_POS %in% POS)
xmdl <- lm(Conc.M ~ SUBTL_POS, conc_red)
anova(xmdl)
summary(xmdl)$r.squared
```

Also significant, explains 32% of the variance.

Get the descriptive values for low SD words:

```{r}
conc_red %>% group_by(SUBTL_POS) %>%
  summarize(M = mean(Conc.M),
            SD = sd(Conc.M),
            N = n())
```

Basically the same pattern as before, but even more pronounced.

Analysis of elastic words. First, check what elastic nouns there are most of, only for those that have at most two POS elasticity and at least 50 instances:

```{r}
conc %>% count(SUBTL_allPOS, sort = TRUE) %>%
  filter(!is.na(SUBTL_allPOS)) %>% 
  filter(str_count(SUBTL_allPOS, '\\.') == 1) %>% 
  filter(n > 50) %>% 
  print(n= Inf)
```

Recode them:

```{r}
conc <- mutate(conc,
               Elasticity = ifelse(SUBTL_allPOS == 'Verb.Noun',
                                   'Noun.Verb', SUBTL_allPOS),
               Elasticity = ifelse(SUBTL_allPOS == 'Adjective.Noun',
                                   'Noun.Adjective', Elasticity),
               Elasticity = ifelse(SUBTL_allPOS == 'Adjective.Adverb',
                                   'Adverb.Adjective', Elasticity),
               Elasticity = ifelse(SUBTL_allPOS == 'Verb.Adjective',
                                   'Adjective.Verb', Elasticity))
```

Extract elastics and non-elastics:

```{r}
elastics <- filter(conc,
                   Elasticity %in% c('Noun.Verb', 'Noun.Adjective',
                                     'Adverb.Adjective', 'Adjective.Verb',
                                     'Noun', 'Adjective', 'Verb', 'Adverb'))
```

Get the averages:

```{r}
elastics %>% group_by(Elasticity) %>% 
  summarize(M = mean(Conc.M),
            SD = sd(Conc.M)) %>% 
  arrange(desc(M))
```

No clear tendency.

Code percentage into continuous variable:

```{r}
conc <- mutate(conc,
               SUBTL_percent = as.numeric(SUBTL_percent))
```

First, get the noun/verb ones, then the others etc.

```{r}
NV <- filter(conc,
             SUBTL_allPOS %in% c('Noun.Verb', 'Verb.Noun'))
AN <- filter(conc,
             SUBTL_allPOS %in% c('Noun.Adjective', 'Adjective.Noun'))
AV <- filter(conc,
             SUBTL_allPOS %in% c('Verb.Adjective', 'Adjective.Verb'))
AA <- filter(conc,
             SUBTL_allPOS %in% c('Adjective.Adverb', 'Adverb.Adjective'))
```

Code them all so that the percentage is with respect to the more concrete category:

```{r}
NV[NV$SUBTL_allPOS == 'Verb.Noun', ]$SUBTL_percent <- 1 -
  NV[NV$SUBTL_allPOS == 'Verb.Noun', ]$SUBTL_percent

AN[AN$SUBTL_allPOS == 'Adjective.Noun', ]$SUBTL_percent <- 1 -
  AN[AN$SUBTL_allPOS == 'Adjective.Noun', ]$SUBTL_percent

AV[AV$SUBTL_allPOS == 'Adjective.Verb', ]$SUBTL_percent <- 1 -
  AV[AV$SUBTL_allPOS == 'Adjective.Verb', ]$SUBTL_percent

AA[AA$SUBTL_allPOS == 'Adverb.Adjective', ]$SUBTL_percent <- 1 -
  AA[AA$SUBTL_allPOS == 'Adverb.Adjective', ]$SUBTL_percent
```

Assess this:

```{r}
with(NV, cor.test(SUBTL_percent, Conc.M))
with(AN, cor.test(SUBTL_percent, Conc.M))
with(AV, cor.test(SUBTL_percent, Conc.M))
with(AA, cor.test(SUBTL_percent, Conc.M))
```

No clear pattern of results. It's not 100% clear what to make of the elasticity results.

## Linguistic factor 2: Number of morphemes:

Descriptive averages:

```{r}
conc %>% group_by(IsCompound) %>% 
  summarize(M = mean(Conc.M),
            SD = sd(Conc.M),
            SD = round(SD, 2))
```

Assess concreteness difference between compounds and non-compounds:

```{r}
with(conc, t.test(Conc.M ~ IsCompound, var.equal = TRUE))
with(conc, cohen.d(Conc.M ~ IsCompound, var.equal = TRUE))
```

Compare this against monomorphemics:

```{r}
conc_comp <- filter(conc,
                    !(IsCompound == 'No compound' & NMorph > 1))
```

Assess concreteness difference between compounds and non-compounds, for monomorphemics as comparison class:

```{r}
# Descriptive stats:

conc_comp %>% group_by(IsCompound) %>% 
  summarize(M = mean(Conc.M),
            SD = sd(Conc.M),
            SD = round(SD, 2))

# Inferential stats:

with(conc_comp, t.test(Conc.M ~ IsCompound, var.equal = TRUE))
with(conc_comp, cohen.d(Conc.M ~ IsCompound, var.equal = TRUE))
```

There are two ways of looking at morphological complexity. One is to look at specific suffixes (Table 1 in the paper).

Get those from the database and exclude "-est", which is inflectional anyway and has such low numbers because of the lemmatization:

```{r}
conc_morph <- filter(conc,
                     IsCompound != 'Compound',
                     NMorph <= 2)
conc_morph <- conc_morph[-which(conc_morph$Suffix == '-est'), ]
```

What about the suffix versus no suffix variable?

```{r}
conc_morph %>% 
  group_by(HasSuffix) %>% 
  summarize(M = mean(Conc.M),
            SD = sd(Conc.M),
            SD = round(SD, 2))

with(conc_morph, t.test(Conc.M ~ HasSuffix, var.equal = TRUE))
with(conc_morph, cohen.d(Conc.M ~ HasSuffix, var.equal = TRUE))
```

Same for low SD dataset (-est does not occur in the low SD dataset):

```{r}
conc_sub %>% 
  group_by(HasSuffix) %>% 
  summarize(M = mean(Conc.M),
            SD = sd(Conc.M),
            SD = round(SD, 2))

with(conc_sub, t.test(Conc.M ~ HasSuffix))
with(conc_sub, cohen.d(Conc.M ~ HasSuffix))
```

Descriptive values for reporting (NA is monomorphemic here):

```{r}
conc_morph %>% group_by(Suffix) %>%
  summarize(M = round(mean(Conc.M), 2),
            SD = round(sd(Conc.M), 2),
            N = n()) %>% 
  arrange(desc(M)) %>% print(n = Inf)
```

Make a model of this (-able is reference category):

```{r}
summary(lm(Conc.M ~ Suffix, data = conc_morph))
summary(lm(Conc.M ~ Suffix, data = conc_morph))$r.squared
```

Significant effect, explains 41% of the variance. The effect is pretty strong.

Do the same for low SD:

```{r}
summary(lm(Conc.M ~ Suffix,
           data = filter(conc_sub,
                         Suffix != '-est',
                         IsCompound != 'Compound')))
summary(lm(Conc.M ~ Suffix,
           data = filter(conc_sub,
                         Suffix != '-est',
                         IsCompound != 'Compound')))$r.squared
```

Even stronger, 59%.

Make a plot of this, including monomorphemics:

```{r, fig.width = 10, fig.height = 6}
# Plot and aesthetics:

conc_suffix <- conc %>% 
  filter(SuffixWithMono != '-est',
         !is.na(SuffixWithMono)) %>% 
  filter(IsCompound != 'Compound') %>% 
  mutate(SuffixWithMono = ifelse(SuffixWithMono == 'monomorphemic',
                                 'mono.', SuffixWithMono)) %>% 
  ggplot(aes(x = reorder(SuffixWithMono, Conc.M), y = Conc.M))

# Geoms:

conc_suffix <- conc_suffix + geom_boxplot(fill = 'steelblue')

# Themes and other properties:
    
conc_suffix <- conc_suffix + 
  labs(x = '', y = 'Concreteness\n') +
  theme_timo +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1,
                                   face = 'bold', size = 14))

# Look and save:
    
conc_suffix

ggsave(plot = conc_suffix, filename = '../figures/suffixes.pdf',
       width = 12, height = 4.5)
```

Make a plot of morpheme count:

```{r, fig.width = 10, fig.height = 6}
# Plot and aesthetics:

conc_NMorph <- conc %>% 
  filter(!is.na(NMorph)) %>% 
  filter(IsCompound != 'Compound') %>% 
  ggplot(aes(x = factor(NMorph), y = Conc.M,
                    fill = factor(NMorph)))

# Geoms:

conc_NMorph <- conc_NMorph + geom_boxplot()

# Themes and other properties:
    
conc_NMorph <- conc_NMorph + 
  labs(x = '\nNumber of morphemes', y = 'Concreteness\n') +
  scale_fill_brewer(palette = 'Greens') +
  theme_timo +
  theme(legend.position = 'none')

# Look and save:
    
conc_NMorph

ggsave(plot = conc_NMorph, filename = '../figures/morpheme_count.pdf',
       width = 8, height = 4.5)
```

Words with fewer morphemes definitely more likely to be abstract (low concreteness)!

Descriptive values for reporting:

```{r}
conc %>% group_by(NMorph) %>%
  filter(IsCompound != 'Compound') %>% 
  summarize(M = round(mean(Conc.M), 2),
            SD = round(sd(Conc.M), 2),
            N = n()) %>%
  filter(!is.na(NMorph))
```

Make a model of this:

```{r}
summary(lm(Conc.M ~ NMorph,
           data = filter(conc, IsCompound != 'Compound')))
summary(lm(Conc.M ~ NMorph,
           data = filter(conc, IsCompound != 'Compound')))$r.squared
```

Significant effect, explains 24% of the variance. The effect is pretty strong. The coefficient says that for each morpheme, the average concreteness scale is predicted to be lower by half a concreteness point (-0.5). So for two morphemes, you tend to loose a complete concreteness point.

Same for low SD:

Make a model of this:

```{r}
summary(lm(Conc.M ~ NMorph,
           data = filter(conc_sub, IsCompound != 'Compound')))
summary(lm(Conc.M ~ NMorph,
           data = filter(conc_sub, IsCompound != 'Compound')))$r.squared
```

Here's an attempt:

```{r}
xmdl <- lm(Conc.M ~ NMorph + NPhon,
           data = filter(conc, IsCompound != 'Compound'))
summary(xmdl)
summary(xmdl)$r.squared

# Check variance inflation factors to assess collinearity:

vif(xmdl)
```

Adding number of phonemes means that the number of morpheme predictor is still significant... although it is slightly weaker. Adding number of phonemes only increases the variance explained by 19%.

Variance inflation factors indicate no collinearity issues.

Check the number of morphemes now for the low SD subset:

```{r}
# Pure model:

xmdl <- lm(Conc.M ~ NMorph, conc_sub)
summary(xmdl)
summary(xmdl)$r.squared

# With word length control:

xmdl <- lm(Conc.M ~ NMorph + NPhon, conc_sub)
summary(xmdl)
summary(xmdl)$r.squared
```

Significant effects, 20% and 22% of the variance explained.

Descriptive values for reporting of the low SD set:

```{r}
conc_sub %>% group_by(NMorph) %>%
  summarize(M = round(mean(Conc.M), 2),
            SD = round(sd(Conc.M), 2),
            N = n()) %>%
  filter(!is.na(NMorph))
```

There's only three six-morphemic words, so basically not even worth mentioning.

## Linguistic factor 3: Countability:

Needs to be at the lemma level. Let's make a table of "major_class" by lemma:

```{r}
BECL_tab <- with(BECL, table(lemma, major_class))
head(BECL_tab) # first six rows
```

Let's get those that are ONLY count or ONLY noun:

```{r}
colnames(BECL_tab)

# Count nouns:

count_only <- BECL_tab[BECL_tab[, 3] >= 1, ]
count_only <- count_only[count_only[, 1] == 0, ]
count_only <- count_only[count_only[, 2] == 0, ] 
count_only <- count_only[count_only[, 4] == 0, ] 

# Mass nouns:

mass_only <- BECL_tab[BECL_tab[, 4] >= 1, ]
mass_only <- mass_only[mass_only[, 1] == 0, ]
mass_only <- mass_only[mass_only[, 2] == 0, ] 
mass_only <- mass_only[mass_only[, 3] == 0, ]

# Extract the actual words:

mass_only <- row.names(mass_only)
count_only <- row.names(count_only)
```

In the "conc" data frame, use these word vectors to label the words accordingly:

```{r}
conc <- mutate(conc, MassCount = ifelse(Word %in% mass_only, 'mass', NA),
               MassCount = ifelse(Word %in% count_only, 'count', MassCount))

# Low SD set:

conc_sub <- mutate(conc_sub, MassCount = ifelse(Word %in% mass_only, 'mass', NA),
               MassCount = ifelse(Word %in% count_only, 'count', MassCount))
```

Check how many there are:

```{r}
table(conc$MassCount)
prop.table(table(conc$MassCount))

# Low SD set:

table(conc_sub$MassCount)
prop.table(table(conc_sub$MassCount))
```

80% count (N = 3699), and 20% mass (N = 918); very similar proportions for the low SD set.

Make a plot of this:

```{r, fig.width = 8, fig.height = 6}
# Plot and aesthetics:

conc_MassCount <- conc %>% filter(!is.na(MassCount)) %>%
  mutate(MassCount = factor(MassCount, levels = c('mass', 'count'))) %>% 
  ggplot(aes(x = MassCount, fill = MassCount, y = Conc.M))

# Geoms:

conc_MassCount <- conc_MassCount +
  geom_boxplot()

# Themes and other properties:

conc_MassCount <- conc_MassCount +
  labs(y = 'Concreteness\n', x = '') + 
  scale_fill_brewer(palette = 'Accent') +
  theme_timo +
  theme(legend.position = 'none')

# Look and save:

conc_MassCount

ggsave(plot = conc_MassCount, filename = '../figures/masscount.pdf',
       width = 6, height = 4.5)
```

Looks like mass is much more abstract. Descriptive values for comparison:

```{r}
conc %>% group_by(MassCount) %>%
  summarize(M = mean(Conc.M),
            SD = sd(Conc.M),
            N = n()) %>%
  filter(!is.na(MassCount))

# Same for low SD:

conc_sub %>% group_by(MassCount) %>%
  summarize(M = mean(Conc.M),
            SD = sd(Conc.M),
            N = n()) %>%
  filter(!is.na(MassCount))
```

Let's make a model of this:

```{r}
with(conc, t.test(Conc.M ~ MassCount, var.equal = TRUE))
with(conc, cohen.d(Conc.M ~ MassCount, var.equal = TRUE))
with(conc, summary(lm(Conc.M ~ MassCount, ))$r.squared)

# Low SD:

with(conc_sub, t.test(Conc.M ~ MassCount, var.equal = TRUE))
with(conc_sub, cohen.d(Conc.M ~ MassCount, var.equal = TRUE))
with(conc_sub, summary(lm(Conc.M ~ MassCount))$r.squared)
```

Yep, and explains 14% of the variance, 19% for the low SD words.

## Factor 4: Etymology

For how many of the words do we have etymological information?

```{r}
sum(!is.na(conc$Etym))
sum(is.na(conc$Etym))
```

Are mass nouns more likely to be French origin?

```{r}
xtab <- with(conc, table(MassCount, EtymSimplified))
chisq.test(xtab)
```

This significance test tests for an uneven distribution of mass/count labels across French/English and other words. We can look at the standardized residuals...

```{r}
chisq.test(xtab)$stdres
```

... which show that French words are over-represented for mass nouns, and English words over-represented for count nouns.

What about morphological complexity? First, descriptive stats:

```{r}
conc %>% group_by(EtymSimplified) %>% 
  summarize(M = mean(NMorph, na.rm = TRUE)) %>% 
  filter(!is.na(EtymSimplified)) %>% 
  arrange(desc(M))
```

Descriptive stats of concreteness:

```{r}
conc %>% group_by(EtymSimplified) %>% 
  summarize(M = mean(Conc.M, na.rm = TRUE),
            SD = sd(Conc.M, na.rm = TRUE)) %>% 
  filter(!is.na(EtymSimplified)) %>% 
  arrange(desc(M))
```

Let's make a Poisson model of this:

```{r}
nmorph_mdl <- glm(NMorph ~ EtymSimplified,
                  data = conc, family = 'poisson')
summary(nmorph_mdl)
anova(nmorph_mdl, test = 'Chisq')
```

Make a model out of this:

```{r}
etym_mdl <- lm(Conc.M ~ EtymSimplified, data = conc)
summary(etym_mdl)

etym_sub_mdl <- lm(Conc.M ~ EtymSimplified, data = conc_sub)
summary(etym_sub_mdl)
```

Make a graph:

```{r, fig.width = 8, fig.height = 6}
# Plot and aesthetics:

etym_p <- conc %>%
  filter(EtymSize == 'large') %>%
  ggplot(aes(x = EtymSimplified, fill = EtymSimplified, y = Conc.M))

# Geoms:

etym_p <- etym_p +
  geom_boxplot()

# Themes and other properties:
  
etym_p <- etym_p +
  labs(x = '', y = 'Concreteness\n') +
  scale_fill_brewer(palette = 'Accent') + 
  theme_timo +
  theme(legend.position = 'none')
  

# Look and save:

etym_p

ggsave(plot = etym_p, filename = '../figures/etymology.pdf',
       width = 6, height = 4.5)

```

## Simultaneous analysis

Put everything into one regression model, with count and mass nouns treated as separate words:

```{r}
conc$SUBTL_POS2 <- conc$SUBTL_POS
conc[which(conc$MassCount == 'mass'), ]$SUBTL_POS2 <- 'mass noun'
conc[which(conc$MassCount == 'count'), ]$SUBTL_POS2 <- 'count noun'
conc_red <- filter(conc,
                   SUBTL_POS2 != 'Noun')
```

Do the same thing for the low-SD set:

```{r}
conc_sub$SUBTL_POS2 <- conc_sub$SUBTL_POS
conc_sub[which(conc_sub$MassCount == 'mass'), ]$SUBTL_POS2 <- 'mass noun'
conc_sub[which(conc_sub$MassCount == 'count'), ]$SUBTL_POS2 <- 'count noun'
conc_sub_red <- filter(conc_sub,
                       SUBTL_POS2 != 'Noun')
```

Get a subset that has only those 

Run a model:

```{r}
full_mdl <- lm(Conc.M ~ SUBTL_POS2 + NMorph + EtymSimplified,
               data = filter(conc_red, IsCompound != 'Compound'))
summary(full_mdl)$r.squared # full model
anova(full_mdl)# significant effects of everything
```

Get the unique contributions against the full model:

```{r}
mdl_no_SUBTL <- lm(Conc.M ~ NMorph + EtymSimplified,
               data = filter(conc_red, IsCompound != 'Compound'))
mdl_no_morph <- lm(Conc.M ~ SUBTL_POS2 + EtymSimplified,
               data = filter(conc_red, IsCompound != 'Compound'))
mdl_no_etym <- lm(Conc.M ~ SUBTL_POS2 + NMorph,
               data = filter(conc_red, IsCompound != 'Compound'))

# r-squareds:

rsq_all <- summary(full_mdl)$r.squared
rsq_all - summary(mdl_no_SUBTL)$r.squared
rsq_all - summary(mdl_no_morph)$r.squared
rsq_all - summary(mdl_no_etym)$r.squared
```

Partial r-squared:

```{r}
rsq.partial(full_mdl)
```

Similar picture.

Run a model on the low-SD data::

```{r}
sub_mdl <- lm(Conc.M ~ SUBTL_POS2 + NMorph + EtymSimplified,
              data = filter(conc_sub_red, IsCompound != 'Compound'))
summary(sub_mdl)$r.squared # full model
anova(sub_mdl)# significant effects of everything
```

Low-SD partial R-squareds:

```{r}
rsq.partial(sub_mdl)
```

Overall, this model explains 36% of the variance in concreteness ratings.

Check variance inflation factors:

```{r}
vif(full_mdl)
```

Absolutely no collinearity worthy of mentioning.

Do the same thing for the low SD words:

```{r}
conc_red <- filter(conc_sub, SUBTL_POS %in% POS)
full_mdl_sub <- lm(Conc.M ~ NMorph + SUBTL_POS + EtymSimplified,
                   data = conc_red)
anova(full_mdl_sub)# significant effects of everything
summary(full_mdl_sub)
```

Explains 27% of the total variance for the low SD model.

## Subset analysis for high-and low concreteness words:

Check whether linguistic factors explain results better for high or low abstractness words:

```{r}
conc_red_highC <- filter(conc_red, Conc.M > median(Conc.M))
conc_red_lowC <- filter(conc_red, Conc.M < median(Conc.M))
```

The corresponding models:

```{r}
full_highC <- lm(Conc.M ~ SUBTL_POS + NMorph + EtymSimplified,
                 data = conc_red_highC)
full_lowC <- lm(Conc.M ~ SUBTL_POS + NMorph + EtymSimplified,
                data = conc_red_lowC)
anova(full_highC)
anova(full_lowC)

summary(full_highC)$r.squared
summary(full_lowC)$r.squared
```


## Calculate unique R-squared for all things:

Loop through predictors and drop them to look at how much variance they contribute. First, create tibbles where there's no missingness:

```{r}
nomiss <- filter(conc,
                 !is.na(SUBTL_POS),
                 !is.na(NMorph),
                 !is.na(EtymSimplified))
nomiss_sub <- filter(conc_sub,
                     !is.na(SUBTL_POS),
                     !is.na(NMorph),
                     !is.na(EtymSimplified))
```

Create a vector with all the predictors:

```{r}
preds <- c('SUBTL_POS', 'NMorph', 'EtymSimplified')
```

Create tables to append the different R-squared values to:

```{r}
rsq <- tibble(preds, RSQ = NA, RSQ_sub = NA) %>% 
  mutate(RSQ = as.numeric(RSQ),
         RSQ_sub = as.numeric(RSQ_sub))
```

Then, loop through all the models and drop one of the terms:

```{r}
for (i in seq_along(preds)) {
  myFormula <- str_c(preds[-i], collapse = ' + ')
  myFormula <- str_c('Conc.M ~ ', myFormula)
  rsq[i, ]$RSQ <- glance(lm(myFormula, data = nomiss))$r.squared
  rsq[i, ]$RSQ_sub <- glance(lm(myFormula, data = nomiss_sub))$r.squared
}
```

Calculate the R-squared differences:

```{r}
rsq <- mutate(rsq,
              RSQ_diff = glance(full_mdl)$r.squared - RSQ,
              RSQ_sub_diff = glance(full_mdl_sub)$r.squared - RSQ_sub)
rsq
```

Visualization of the R-squared values:

```{r, fig.width = 12, fig.height = 6}
rsq %>%
  ggplot(aes(x = preds, y = RSQ_sub_diff, fill = preds)) +
  geom_bar(stat = 'identity') +
  theme_timo +
  scale_fill_brewer(palette = 'Accent') +
  ggtitle('R-Squared by predictor')
ggsave('../figures/r_squared.pdf', width = 8, height = 4.5)
```

## Do this analysis for nouns only:

Here we can add mass/count and specificity as predictors:

```{r}
nouns <- filter(conc, SUBTL_POS == 'Noun')
noun_mdl <- lm(Conc.M ~ NMorph + MassCount + EtymSimplified, data = nouns)
anova(noun_mdl)
summary(noun_mdl)
```

Same for reduced data frame:

```{r}
noun_sub <- filter(conc_sub, SUBTL_POS == 'Noun')
noun_sub_mdl <- lm(Conc.M ~ NMorph + MassCount + EtymSimplified,
                   data = noun_sub)
anova(noun_sub_mdl)
summary(noun_sub_mdl)
```

Get data frames:

```{r}
nomiss_noun <- filter(conc,
                 SUBTL_POS == 'Noun',
                 !is.na(NMorph),
                 !is.na(MassCount),
                 !is.na(EtymSimplified))
nomiss_noun_sub <- filter(conc_sub,
                 SUBTL_POS == 'Noun',
                 !is.na(NMorph),
                 !is.na(MassCount),
                 !is.na(EtymSimplified))
```

Create a vector with all the predictors:

```{r}
preds <- c('MassCount', 'NMorph', 'EtymSimplified')
```

Create tables to append the different R-squared values to:

```{r}
rsq_nouns <- tibble(preds, RSQ = NA, RSQ_sub = NA) %>% 
  mutate(RSQ = as.numeric(RSQ),
         RSQ_sub = as.numeric(RSQ_sub))
```

Then, loop through all the models and drop one of the terms:

```{r}
for (i in seq_along(preds)) {
  myFormula <- str_c(preds[-i], collapse = ' + ')
  myFormula <- str_c('Conc.M ~ ', myFormula)
  rsq_nouns[i, ]$RSQ <- glance(lm(myFormula, data = nomiss_noun))$r.squared
  rsq_nouns[i, ]$RSQ_sub <- glance(lm(myFormula, data = nomiss_noun_sub))$r.squared
}
```

Calculate the R-squared differences:

```{r}
rsq_nouns <- mutate(rsq_nouns,
              RSQ_diff = glance(noun_mdl)$r.squared - RSQ,
              RSQ_sub_diff = glance(noun_sub_mdl)$r.squared - RSQ_sub)
rsq_nouns
```

Visualization of the R-squared values:

```{r, fig.width = 12, fig.height = 6}
rsq_nouns %>%
  ggplot(aes(x = preds, y = RSQ_sub_diff, fill = preds)) +
  geom_bar(stat = 'identity') +
  theme_timo +
  scale_fill_brewer(palette = 'Accent') +
  ggtitle('R-Squared by predictor')
```

This completes this analysis.
